---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


## Installing the BiBitR Package


```{r, eval=FALSE,echo=TRUE}
install.packages("devtools") # If not yet installed on your R Version
devtools::install_github("hadley/devtools") # Only run this if your currently installed 
                                            # devtools version is <= 1.12 (recursive dependencies bug)

devtools::install_github("ewouddt/BiBitR")
```

Should the installation of `BiBitR` or `devtools::install_github("hadley/devtools")` throw an error, please install the dependencies manually, then `BiBitR`:
```{r, eval=FALSE,echo=TRUE}
install.packages(c("flexclust","biclust"))
devtools::install_github("ewouddt/BiBitR")
```

## Details

`BiBitR` is a simple R wrapper which directly calls the original Java code for applying the BiBit algorithm after which the output is transformed to a `Biclust` S4 class object.
The original Java code can be found at [http://eps.upo.es/bigs/BiBit.html](http://eps.upo.es/bigs/BiBit.html) by Domingo S. Rodriguez-Baena, Antonia J. Perez-Pulido and Jesus S. Aguilar-Ruiz.

More details about the **BiBit** algorithm can be found in:

* [Domingo S. Rodriguez-Baena, Antonia J. Perez-Pulido and Jesus S. Aguilar-Ruiz (2011), "A biclustering algorithm for extracting bit-patterns from binary datasets", *Bioinformatics*](http://bioinformatics.oxfordjournals.org/content/early/2011/08/08/bioinformatics.btr464.abstract).


The `bibit` function uses the original Java code directly (with the intended input and output). Because the Java code was not refactored, the `rJava` package could not be used.

The `bibit` function does the following:

1. Convert R matrix to a `.arff` output file.
2. Use the `.arff` file as input for the Java code which is called by `system()`.
3. The outputted `.txt` file from the Java BiBit algorithm is read in and transformed to a `Biclust` object.

Because of this procedure, there is a chance of *overhead* when applying the algorithm on large datasets. Make sure your machine has enough RAM available when applying to big data.


## Return Value

A Biclust S4 Class object.

## Example

```{r, eval=FALSE,echo=TRUE}
library(BiBitR)

data <- matrix(sample(c(0,1),100*100,replace=TRUE,prob=c(0.9,0.1)),nrow=100,ncol=100)
data[1:10,1:10] <- 1 # BC1
data[11:20,11:20] <- 1 # BC2
data[21:30,21:30] <- 1 # BC3
data <- data[sample(1:nrow(data),nrow(data)),sample(1:ncol(data),ncol(data))]

result <- bibit(data,minr=5,minc=5)
result

MaxBC(result)
```

